Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-09-29 paper.tex
intro on performance

@@

        \section{Modularity}\label{mod}
            \tristan{Since this section is about interfaces, I would
            rather call it ``Interfaces" or ``Tool interfaces".}
            Often times, users will want to integrate existing tools, in 
            addition to custom scripts, into their workflows. As workflow 
            engines, with the exception of Make, are typically not written in 
            Shell, it is necessary to create interfaces to these tools. These
            interfaces may range from simple calls to the command-line, to more 
            complex ones that track inputs, outputs and other parameters; ensure
            that all parameters are specified in the correct order and that the
            parameter and file types are all correct; capture all of the 
            standard input/output; and create folders to encapsulate task inputs
            and outputs. Workflow engines typically rely on more complex types 
            of interfaces in order to ensure correct functioning of the pipeline
            . As these interfaces can be rather complex, it is necessary for the 
            workflow engine to provide an intuitive way for users to create 
            interfaces for their custom tools.

@@

            OpenMOLE has the ability to encapsulate UDF Scala tasks and CLI. 
            In the case of CLI tasks, it is also possible to containerize or 
            encapsulate with CARE \note{cite} to make it portable.

@@

@@

        \section{Workflow sharing}\label{sharing}
    \chapter{Performance}\label{performance}
        \section{Filesystems}
            scientific: shared FS such as Lustre, NFS, S3
            big data: mapreduce - HDFS, spark and dask shared or distributed
        \section{Scheduling}
            scientific: SLURM, PBS, SGE, HTCondor, Mesos (sometimes)
            big data: mapreduce - local or YARN, Spark - standalone, YARN or 
            Mesos, dask: any (easy to extend to any scheduler)
        \section{Performance enhancement strategies}
            scientific: workflow checkpointing (looking at file hashes)
            big data: data locality, in memory computing, lazy evaluation
    \chapter{Portability}\label{portability}
        java or python backend (most) typically platform independent (some 
        exceptions)
        all tools used in the pipeline must be installed on system
        Taverna - exception as relies on web services
        \section{Containers}
            use of containers increase portability. none of the scientific 
            pipelines can be containerized and communicate with the scheduler 
            outside of the container
            big data frameworks can
    \chapter{Reproducibility}\label{reproducibility}
        \section{Provenance Tracking}
        big data frameworks: limited to lineage
        \section{containerization}
    \chapter{Discussion/Conclusion}

>>>>

            \subsection{Data structures}
                \subsubsection{Scientific workflows}
                \subsubsection{Big Data: MapReduce}

@@

        \section{Tool interfaces}\label{mod}
            \tristan{Since this section is about interfaces, I would
            rather call it ``Interfaces" or ``Tool interfaces".}
            Often times, users will want to integrate existing tools, in 
            addition to custom scripts, into their workflows. As workflow 
            engines, with the exception of Make, are typically not written in 
            Shell, it is necessary to create interfaces to these tools. These
            interfaces may range from simple calls to the command-line, to more 
            complex ones that track inputs, outputs and other parameters; ensure
            that all parameters are specified in the correct order and that the
            parameter and file types are all correct; capture all of the 
            standard input/output; and create folders to encapsulate task inputs
            and outputs. Workflow engines typically rely on more complex types 
            of interfaces in order to ensure correct functioning of the pipeline
            . As these interfaces can be rather complex, it is necessary for the 
            workflow engine to provide an intuitive way for users to create 
            interfaces for their custom tools.

@@

            OpenMOLE has the ability to encapsulate UDF Scala scripts and CLI. 
            In the case of CLI tasks, it is also possible to containerize or 
            encapsulate with CARE \note{cite} to make it portable. OpenMOLE
            defines several classes to allow for the execution of different 
            types of tools. Like other framworks, it is vital to describe the
            inputs and outputs of the Tasks such that the execution of tasks 
            can be performed.

@@

            Unlike scientific workflow engines, BigData frameworks do not 
            provide default interfaces for the execution of tools written in 
            languages different from their own. Users defining their workflows
            in Dask or the Spark and MapReduce using the Python API integrate
            Python UDFs in their workflow. While Dask users can reference their 
            UDF of Python function directly in their workflow, Spark users
            must define a lambda function within an action or transformation 
            that executes their UDF or library on the data. Similarly to Spark,
            functions referenced in MapReduce must be called from within a Map
            or Reduce task. MapReduce can additionally reference Java 
            code and Spark can call on code also written in Scala, Java and R.
            \note{verify if it's possible to mix-and-match different libraries
            ...don't think so}

            For all workflow engines that do not, by default, interface with 
            command-line tools or containers, there are libraries that can be
            easily integrated to achieve that. Boutiques, for instance, is a 
            Python library that can execute command-line or containerized tools
            defined within a JSON schema. Furthermore, the library's richness
            compete with Galaxy's XML. Python-based BigData frameworks can 
            easily call on this library from within the workflow in order to
            communicate with the tools.

            \subsection{input and output definitions}
                TBD

@@

        \section{Workflow sharing}\label{sharing}

    \chapter{Performance}\label{performance}
        Neuroimaging frameworks are notoriously known to be compute-intensive.
        The tools and their respective workflows needed to incorporate 
        parallelism in order to save on computing time. Maintaining good 
        performance has become especially critical now that BigData has been
        introduced into the mix. The dichotomy remains between scientific 
        workflows and BigData workflows. While scientific workflows have made
        advances towards more performant engines, they still lack the 
        strategies BigData frameworks employ to reduce data transfers.

        Another notable aspect to consider is that both types of engines 
        (scientific and BigData) are tightly coupled to the infrastructure
        they were intended to be used on. That is, as scientific workflow 
        try to remain efficient in their handling of compute resources, as such
        resources were previously scarce. In contrast, BigData frameworks 
        make efficient use of commodity clusters, where data transfers are 
        burdensome, particularly when faced with BigData.

        As some researchers are trying to shift towards the use of commodity 
        infrastructure while others are trying to process BigData on HPC, the 
        ideal workflow engine must run efficiently regardless of 
        infrastructure available. Otherwise, researchers are either left with
        workflows running inefficiently on resources or having to learn 
        multiple frameworks and rewrite their workflows depending on the 
        infrastructure selected.

        In this Chapter, we will compare scientific
        workflows to BigData frameworks on the basis of performance. 
        Section~\ref{fs} will look at what filesystems are required by the 
        engines. This is important because data movement of any kind is costly,
        and may be worse depending on the filesystems available. 
        Section~\ref{sched} will provide an overview of the schedulers 
        available and their compatibility with the workflow engines. 
        Section~\ref{other} will discuss other performance improvement 
        strategies employed by the workflow engines.

        \section{Filesystems}\label{fs}

            Scientific workflows historically relied on I/O to transfer data
            between nodes. As scientific workflows are usually executed on
            HPC infrastructure, this meant writing to a filesystem such as 
            Lustre

            \subsection{Lustre}
            \subsection{Hadoop Distributed File System}
            \subsection{Amazon S3}

            scientific: shared FS such as Lustre, NFS, S3
            big data: mapreduce - HDFS, spark and dask shared or distributed
        \section{Scheduling}\label{sched}
            scientific: SLURM, PBS, SGE, HTCondor, Mesos (sometimes)
            big data: mapreduce - local or YARN, Spark - standalone, YARN or 
            Mesos, dask: any (easy to extend to any scheduler)
        \section{Performance enhancement strategies}\label{other}
            scientific: workflow checkpointing (looking at file hashes)
            big data: data locality, in memory computing, lazy evaluation
    \chapter{Portability}\label{portability}
        java or python backend (most) typically platform independent (some 
        exceptions)
        all tools used in the pipeline must be installed on system
        Taverna - exception as relies on web services
        \section{Containers}
            use of containers increase portability. none of the scientific 
            pipelines can be containerized and communicate with the scheduler 
            outside of the container
            big data frameworks can
    \chapter{Reproducibility}\label{reproducibility}
        \section{Provenance Tracking}
        big data frameworks: limited to lineage
        \section{containerization}
    \chapter{Discussion/Conclusion}
